\magnification=1200
\centerline{\bf ENIAC Simulator Configuration Notation Reference}
\vskip 12pt
\noindent
{\bf Initiating Unit}

Terminals

{\advance\leftskip 0.5in
\item{\tt Ci$n$:} Selective clear input ($1\le n\le 6$)
\item{\tt Co$n$:} Selective clear ouput ($1\le n\le 6$)
\item{\tt Ri:} Reader input
\item{\tt Ro:} Reader output
\item{\tt Rl:} Reader interlock
\item{\tt Pi:} Printer/punch input
\item{\tt Po:} Printer/punch output
\item{\tt Pl:} Printer/punch interlock
\item{\tt Io:} Initiating pulse output

}

Buttons

{\advance\leftskip 0.5in
\item{\tt c:} Initial clear
\item{\tt i:} Initiating pulse
\item{\tt r:} Reader start switch

}

Examples

{\advance\leftskip 0.5in
\item{1.} Press the initiating pulse button: {\tt b i}
\item{2.} Plug a program jumper from the initiating pulse
output of the initiating unit to program trunk 1, line 2:  {\tt p i.Io 1-2}

}

Notes

{\advance\leftskip 0.5in
\item{1.} Although the clear, reader, printer, and initiate mnemonics
are show as upper case here and on most of the original diagrams,
either upper or lower case are accecpted by the simulator.
\item{2.} The terminals for {\tt Rs} and {\tt Is} are not implemented as these
are assumed connected to the hand-held control unit.
\item{3.} The Start, Stop, and Door Shut switches are not implemented.

}

\vskip 12pt
\noindent
{\bf Cycling Unit}

Switch

{\advance\leftskip 0.5in
\item{\tt op:} Cycling unit operating mode: 1a (1 add), 1p (1 pulse), or co (continuous)

}

Button

{\advance\leftskip 0.5in
\item{\tt p:} 1 pulse \& 1 add time button

}

Examples

{\advance\leftskip 0.5in
\item{1.} Set the cycling unit to 1 add mode:  {\tt s cy.op 1a}
\item{2.} Single step in 1a or 1p modes: {\tt b p}

}

Notes

{\advance\leftskip 0.5in
\item{1.} The PA, Ext, 1A, and Cont terminals are unimplemented.
\item{2.} The heater switch is unimplemented.
\item{3.} The oscilloscope and oscilloscope input switch are unimplemented.
\item{4.} Values for the operation switch may be specified in either upper
or lower case.

}

\vfill\eject
\vskip 12pt
\noindent
{\bf Constant Transmitter}

Terminals

{\advance\leftskip 0.75in
\item{\tt o:} Digit output terminal
\item{\tt $n$i:} Program input terminals ($1\le n\le 30$)
\item{\tt $n$o:} Program output terminals ($1\le n\le 30$)

}

Switches

{\advance\leftskip 0.75in
\item{\tt s$n$:} Constant selector ($1\le n\le 30$): {\tt Al}, {\tt Ar}, {\tt Alr}, ..., {\tt Kl}, {\tt Kr}, {\tt Klr}
\item{\tt jl, jr:}
\item{\tt kl, kr:} PM set: {\tt P} or {\tt M}
\item{\tt j$n$, k$n$:} Constant set ($1\le n\le 10$): 0--9

}

\vskip 12pt
\noindent
{\bf Printer}

Switches

{\advance\leftskip 0.75in
\item{$n$--$m$:} Coupling ($1\le n\le 8$, $m=n+1$): {\tt 0} or {\tt C}
\item{$n$:} Printing ($1\le n\le 16$): {\tt O} or {\tt P}

}

\vskip 12pt
\noindent
{\bf Accumulator}

Terminals

{\advance\leftskip 0.5in
\item{$\alpha$--$\epsilon$:} Input data terminals
\item{\tt A:} Additive output data terminal
\item{\tt S:} Subtractive output data terminal
\item{\tt $n$i:} Program input terminal ($1\le n\le 12$)
\item{\tt $n$o:} Program output terminal ($5\le n\le 12$)
\item{\tt I$sn$:} Accumulator interconnect terminals
($s\in\lbrace {\tt l, r}\rbrace$, $n\in\lbrace 1, 2\rbrace$)
\item{\tt lb:} Special notation for a load block on interconnects

}

Switches

{\advance\leftskip 0.5in
\item{\tt op$n$:} Operation select ($1\le n\le 12$): $\alpha$--$\epsilon$, {\tt 0}, {\tt A}, {\tt AS}, or {\tt S}
\item{\tt rp$n$:} Repeat ($5\le n\le 12$): 1--9
\item{\tt cc$n$:} Clear correct ($1\le n\le 12$): {\tt 0} or {\tt C}
\item{\tt sf:} Significant figures: 0--10
\item{\tt sc:} Selective clear: {\tt 0} or {\tt SC}

}

Examples

{\advance\leftskip 0.5in
\item{1.} Plug jumper from data trunk 3 to the gamma input on accumulator 12: {\tt p 3 a12.$\gamma$}
\item{2.} Set program 5 on accumulator 9 to output additively: {\tt s a9.op5 A}
\item{3.} Set accumulator 2 program 7 repeat to 4: {\tt s a2.rp7 4}

}

Notes

{\advance\leftskip 0.5in
\item{1.} Carry clear and selective clear switch settings may be specified in
either upper or lower case.
\item{2.} The input data terminals and input operation switch settings may be
specified with either  a Unicode Greek letter, the lower case spelled-out name of the Greek
letter, or the corresponding lower case Roman letter.

}

\vfill\eject
\noindent
{\bf High Speed Multiplier}

Terminals

{\advance\leftskip 1in
\item{\tt lhppI:}
\item{\tt lhppII:}
\item{\tt rhppI:}
\item{\tt rhppII:} Partial product data terminals
\item{{\tt R$\alpha$}--{\tt R$\epsilon$:}} Multiplier accumulator program pulse outputs
\item{{\tt D$\alpha$}--{\tt D$\epsilon$:}} Multiplicand accumulator program pulse outputs
\item{\tt A, S, AS:}
\item{\tt AC, SC, ASC:} Product disposal terminals
\item{\tt RS, DS, F:} Internal operating terminals
\item{\tt $n$i:} Program input terminals ($1\le n \le 24$)
\item{\tt $n$o:} Program output terminals ($1\le n \le 24$)

}


\vskip 6pt
Switches

{\advance\leftskip 1in
\item{\tt ieracc$n$:} Multiplier accumulator receive switch ($1\le n\le 24$): $\alpha$--$\epsilon$ or {\tt 0}
\item{\tt icandacc$n$:} Multiplicand accumulator receive switch
($1\le n\le 24$): $\alpha$--$\epsilon$ or {\tt 0}
\item{\tt sf$n$:} Significant figures switch ($1\le n\le 24$): 0, 2--10
\item{\tt place$n$:} Multiplier places switch ($1\le n\le 24$): 2--10
\item{\tt prod$n$:} Product disposal switch ($1\le n\le 24$): {\tt A}, {\tt S}, {\tt AS}, {\tt 0},
{\tt AC}, {\tt SC}, or {\tt ASC}
\item{\tt iercl$n$:} Multiplier accumulator clear switch ($1\le n\le 24$): {\tt 0} or {\tt C}
\item{\tt icandcl$n$:} Multiplicand accumulator clear switch ($1\le n\le 24$): {\tt 0} or {\tt C}

}

\vskip 6pt
Typical Multiplier Setup

\halign{\qquad\qquad\tt #\quad & \tt #\quad & \tt #\quad & \tt #\cr
{\rm Multiplier} & {\rm Multiplicand} & {\rm LHPP} & {\rm RHPP} \cr
p m.R$\alpha$ 9-1 & p m.D$\alpha$ 9-2 & p 9-4 a11.2i & p m.A 9-3\cr
p 9-1 a9.1i & p 9-2 a10.1i & p m.F 9-6 & p 9-3 a13.3i\cr
p m.RS 9-4 & p m.DS 9-5 & p 9-6 a11.1i & p 9-5 a13.2i\cr
p 9-4 a9.6i & p 9-5 a10.6i & p 7 a11.$\beta$ & p 9-6 a13.1i\cr
p a9.S 7 & p a10.S 6 & p a11.A 6 & p 6 a13.$\beta$ \cr
s m.ieracc1 $\alpha$ & s m.icandacc1 $\alpha$ & p m.lhppI 8 & p m.rhppI 9\cr
s m.iercl1 C & s micandcl1 C & p 8 a11.$\alpha$ & p 9 a13.$\alpha$\cr
s a9.op1 $\alpha$ & s a10.op1 $\alpha$ & s a11.op1 A & s a13.op1 $\beta$\cr
s a9.op6 S & s a10.op6 S & s a11.cc1 C & s a13.op2 $\beta$\cr
s a9.rp6 1 & s a10.rp6 1 & s a11.op2 $\beta$ & s a13.op3 A\cr
& & s m.sf1 8 & \cr
& & s m.place1 2 & \cr
& & s m.prod1 A & \cr
}
\vfill\eject
\noindent
{\bf Divider \& Square Rooter}

Terminals

{\advance\leftskip 0.5in
\item{\tt $n$i:} Program input terminals ($1\le n\le 8$)
\item{\tt $n$o:} Program output terminals ($1\le n\le 8$)
\item{\tt $n$l:} Program interlock terminals ($1\le n\le 8$)
\item{\tt ans:} Digit answer output terminal

}

Adapters (specified as switches)

{\advance\leftskip 0.5in
\item{\tt da:} Divider adapter set: {\tt A}, {\tt B}, or {\tt C}
\item{\tt ra:} Square Rooter adapter set: {\tt A}, {\tt B}, or {\tt C}

}

Switches

{\advance\leftskip 0.5in
\item{\tt nr$n$:} Numerator accumulator receive ($1\le n\le 8$): $\alpha$, $\beta$, or {\tt 0}
\item{\tt nc$n$:} Numerator accumulator clear ($1\le n\le 8$): {\tt 0} or {\tt C}
\item{\tt dr$n$:} Denominator-square-root accumulator receive ($1\le n\le 8$): $\alpha$, $\beta$, or {\tt 0}
\item{\tt dc$n$:} Denominator-square-root accumulator clear ($1\le n\le 8$): {\tt 0} or {\tt C}
\item{\tt pl$n$:} Places ($1\le n\le 8$): {\tt D4}, {\tt D7}--{\tt 10}, {\tt S4}, or {\tt S7}--{\tt 10}
\item{\tt ro$n$:} Round-off ($1\le n\le 8$): {\tt RO} or {\tt NRO}
\item{\tt an$n$:} Answer disposal ($1\le n\le 8$): 1--4, or {\tt OFF}
\item{\tt il$n$:} Interlock ($1\le n\le 8$): {\tt I} or {\tt NI}

}

Basic Setup

\halign{\qquad\qquad\tt #\qquad & \tt #\hfil\cr
{\rm Divider} & {\rm Square Rooter}\cr
p d.ans 8 & p d.ans 8\cr
p 8 a2.$\alpha$ & p 8 a2.$\alpha$\cr
p 8 a5.$\gamma$ & p 8 a5.$\gamma$\cr
p a5.A 9 & p a5.A 9\cr
p a5.S 9 & p a5.S 9\cr
p a3.A 9 & p a3.A 9\cr
p a7.A 9 & p a7.A 9\cr
p 9 a3.$\gamma$ & p 9 a3.$\gamma$\cr
p 9 ad.s.1.1 & p 9 ad.s.1.1\cr
p ad.s.1.1 a7.$\alpha$ & p ad.s.1.1 a7.$\alpha$\cr
s d.nr1 $\alpha$ & s d.nr1 $\alpha$\cr
s d.nc1 C & s d.nc1 C\cr
s d.dr1 $\alpha$ & s d.dr1 0\cr
s d.dc1 C & s d.dc1 0\cr
s d.pl1 D10 & s d.pl1 R8\cr
s d.ro1 RO & s d.ro1 RO\cr
s d.an1 1 & s d.an1 4\cr
s d.il1 NI & s d.il1 NI\cr
s d.da A & s d.ra A\cr
}

Notes

{\advance\leftskip 0.5in
\item{1.} Only 10-digit numerators, denominators, and radicands are supported.
\item{2.} Accumulator 2 is assumed for the quotient, Accumulator 3 for the numerator or radicand,
Accumulator 5 for the denominator or square root, and Accumulator 7 for the shift accumulator.

}

\vfill\eject
\noindent
{\bf Function Tables}

Terminals

{\advance\leftskip 0.75in
\item{\tt arg:} Argument input termal
\item{\tt A, B:} Function output terminals
\item{\tt NC, C:} Argument reception NC and C program pulse output terminals
\item{\tt $n$i:} Program input terminals ($1\le n\le 11$)
\item{\tt $n$o:} Program output terminals ($1\le n\le 11$)

}

Switches

{\advance\leftskip 0.75in
\item{\tt op$n$:} Operation ($1\le n\le 11$): {\tt A-2}, {\tt A-1}, ..., {\tt A+2}, {\tt S-2}, ..., {\tt S+2}
\item{\tt cl$n$:} Argument reception ($1\le n\le 11$): {\tt 0}, {\tt C}, or {\tt NC}
\item{\tt rp$n$:} Operation repeat ($1\le n\le 11$): 1--9
\item{\tt mpm$n$:} Master PM switch ($1\le n\le 2$): {\tt P} or {\tt M}
\item{\tt A$n$C:}
\item{\tt B$n$C:} Constant digit ($1\le n\le 4$): 0--9, {\tt PM1}, or {\tt PM2}
\item{\tt A$n$D:}
\item{\tt B$n$D:} Digit delete ($1\le n\le 4$): {\tt O} or {\tt D}
\item{\tt A$n$S:}
\item{\tt B$n$S:} Subtract pulse ($4\le n\le 10$): {\tt 0} or {\tt S}
\item{\tt RA$r$L$d$:}
\item{\tt RB$r$L$d$:} Switch panel digit ($-2\le r\le 101$, $1\le d\le 6$): 0--9
\item{\tt RA$r$S:}
\item{\tt RB$r$S:} Switch panel sign ($-2\le r\le 101$): {\tt P} or {\tt M}
\item{\tt Ninep:} Adapter PX-4-119 (c.f. Sec 7.4): Ninep or Cpp

}

\vskip 12pt
\noindent
{\bf Master Programmer}

Terminals

{\advance\leftskip 1in
\item{\tt $n$di:} Decade direct input ($1\le n\le 20$)
\item{{\tt Adi}--{\tt Kdi:}} Stepper direct input
\item{{\tt Ai}--{\tt Ki:}} Stepper input
\item{{\tt Acdi}--{\tt Kcdi:}} Stepper clear input
\item{{\tt A$n$o}--{\tt K$n$o:}} Stepper output ($1\le n\le 6$)

}

Switches

{\advance\leftskip 1in
\item{\tt a$n$:} Decade associator switch ($n\in \lbrace 2, 4, 8, 10, 12, 14, 18, 20\rbrace$): {\tt A}--{\tt K}
\item{\tt d$n$s$m$:} Decade switch ($1\le n\le 20$, $1\le m\le 6$): 0--9
\item{{\tt cA}--{\tt cK:}} Stepper clear: 1--6

}

\vskip 12pt
\noindent
{\bf Adapters}

Types

{\advance\leftskip 0.5in
\item{\tt dp:} Digit program pulse, e.g. {\tt ad.dp.1.11} for sign
\item{\tt s:} Shift, $>0$ left, $<0$ right
\item{\tt d:} Deleter, e.g. {\tt ad.d.1.7} for xxxxxxxxxx $\to$ xxxxxxx000,
{\tt ad.d.2.-7} for xxxxxxxxxx $\to$ 0000000xxx
\item{\tt sd:} Special digit (used in A.G. Chapter 7)

}

\vfill\eject\end
